---
title: "Introduction to reasytries"
output: rmarkdown::html_vignette
description: >
  Start here if this is your first time using reasytrie. This document gives you
  a comprehensive introduction all the APIs available in the package.
vignette: >
  %\VignetteIndexEntry{Introduction to reasytries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

To use `reasytrie`, load the library like below:

```{r setup}
library(reasytries)
```

At a high level, `reasytrie` implements a trie (aka. prefix tree) data structure
that allows you to search for a word in a dictionary. The time complexity for
the search is $O(m)$, where $m$ is the length of the word to search.

The package provides you the following 4 operations you can do with a trie:

* Word insertion
* Word deletion
* Word complete search
* Word prefix search

This document introduces you to all the operations listed above, by giving you
a complete example involving all these operations.

## Data

For demonstration purposes, we will start with a (incomplete) list of fruit
names from scratch:

```{r}
fruit_names_dictionary = list(
  "apple", "apricot", "avocado", "banana", "berry", "cantaloupe", "cherry", 
  "citron", "citrus", "coconut", "grape", "guava", "kiwi", "lemon", "lime",
  "mango", "melon", "mulberry", "nectarine", "orange", "papaya", "peach",
  "pear", "pineapple", "plum", "prune", "raisin", "raspberry", "tangerine"
)
```

Note that we'll refer to this list of names as "dictionary". This is a list you
will define based on your use case. For example, if your are building a word
search suggestion system, the dictionary will be all the possible suggestion.

## Create a trie

The first step of using a trie is to build it with your dictionary. To do so,
first you need to call `trie_create()` to initialize an empty trie:

```{r}
trie <- trie_create()
```

<mark>
Pay special attention that: the trie created here is a **mutable** structure.
All operations mentioned below modifies the trie in place.
</mark>

Next, you want to add all of your words in yout dictionary into the trie by
calling `trie_add()` for each of the words:

```{r}
for (fruit_name in fruit_names_dictionary) {
  trie_add(trie, fruit_name)
}
```

## Search a trie

The most common use case for a trie is to check if a word is in a trie or what
words start with a prefix. This is implemented in the package as functions
`trie_contain()` and `trie_find_prefix()`.

If you would like to check if a word is in a trie, you can use `trie_contain()`.
`trie_contain()` returns a logical value indicating if a word is in a trie,
like below:

```{r}
# banana is a word in the original dictionary
trie_contain(trie, "banana")

# durian is NOT a word in the original dictionary
trie_contain(trie, "durian")
```

If you would like to get all words starts with a prefix, you can use
`trie_find_prefix()`:

```{r}
# find all words starts with "ap"
trie_find_prefix(trie, "ap")

# find all words starts with "m"
trie_find_prefix(trie, "m")

# find all words starts with "dur" (should be empty)
trie_find_prefix(trie, "dur")
```

## Delete words in a trie

After the creation of a trie, sometimes you might want to remove a word from
the trie if it is no longer valid in your context. You can achieve so with
`trie_delete()`:

```{r}
# Our trie originally contains "apple", so trie_contain should return TRUE at 
#   this point
trie_contain(trie, "apple")

# Now, let's remove "apple" from our trie.
# Should return true indicating deletion success.
trie_delete(trie, "apple")

# Now, trie_contain should return FALSE as "apple" has been removed.
trie_contain(trie, "apple")

# finding words starts with "ap" should no longer gives us "apple" in the result
trie_find_prefix(trie, "ap")
```

On the other hand, deletion will fail if you are trying to remove a word not in
the trie:
```{r}
# Should return FALSE indicating deletion failure.
trie_delete(trie, "durian")
```
